DESING AND SIMULATION SOFTWARE FOR RHODOTRON TYPE ELECTRON ACCELERATORS

### 2023-12-03 subsections/simulation/intermediate_versions/intermediate.tex
calistay puısh

\subsection{$L_{out}$ Optimization For Single \e} \label{sec:lout_sweep}
Initial step of improving \textit{POC} towards \textit{Rhodotron Simulation} was to implement \textit{$L_{out}$ optimizations} to help optimazing magnet designs, as discussed in \fromsec{parameter_sweep}.
First approach was to hang the \e outside of the cavity for $t_{out} = L_{out}/v$, then inject it back to the cavity with reversed $\vecthreeBF{v}$. One can sweep the $t_{out}$ parameter to find the optimal value.
This simple implementation can be found in \fromfig{lout_opt_single_e} of \fromapp{intermediate_codes}.

@@

\subsection{$\phi_{lag}$ Optimization For Bunches} \label{sec:philag_sweep}
After successfully accelerating single \e, particle bunches were implemented to approximate a real \e gun. 
They were modeled as $N$ electrons fired from an \e gun at even time intervals. This approach was taken because the amount of time gun fires, defined as \textit{Gun Active Time, $t_g$}, is a crucial part of pulsed \e gun design.
\clearpage
Addition of bunches would immediately proven useful when finding optimal phase lag for bunch entry.
$\phi_{lag}$ for a bunch was defined as the RF phase when the first \e of the bunch has entered the cavity and it defines the starting time of the current pass.
To use the \textit{parameter sweep} method, as used in \textit{$L_{out}$ optimization}, relevant bunch characteristics are defined as
\vspace{-10pt}\begin{itemize}
    \item $\mu E$: Average energy,
    \item $E_{rms}$: Root mean square of energy,
    \item $R_{rms}$: Root mean square of \e positions.
\end{itemize}

Optimal $\phi_{lag}$ would produce maximum $\mu E$ while minimizing $E_{rms}$ \& $R_{rms}$. For the first pass, 
$E_{rms}$ and $R_{rms}$ would be vaguely dependent of each other; therefore, initial implementation of $\phi_{lag}$ sweep was based on minimizing $E_{rms}$ during simulation (\textit{see \fromfig{phlag_opt}}). 
For $\mu E$ considerations, data from the software would be analyzed either manually or by using external tools such as \textit{ROOT framework}. 
\vspace{20pt}
\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
        int phase_opt(int phase_sweep_range){
            double minrms = 1;
            int opt_phase;
            for(int RFphase = -phase_sweep_range; RFphase <= phase_sweep_range; RFphase++){
              Bunch bunch1(RFphase);
              double t1 = 0;
              bunch1.bunch_gecis_t(t1);
              bunch1.reset_pos();

              if( bunch1.E_rms() < minrms ){
                minrms = bunch1.E_rms();
                opt_phase = RFphase;
              }
            }
            return opt_phase;
        }
    \end{minted}
    \vspace{20pt}
    \caption{$\phi_{lag}$ Optimization For Initial Bunch Design.}
    \label{fig:phlag_opt}
\end{figure}
\clearpage
Since $\phi_{lag}$ is relatively easy to change after production, another version of \fromfig{phlag_opt} that was modified for given magnet design parameters was also implemented (\textit{see \fromfig{phlag_opt_n_pass} in \fromapp{intermediate_codes}}). 
This version can be useful for optimizing $\phi_{lag}$ in case of production issues in magnets.

@@

\subsection{Simulation in 3D}
After successfully implementing \eE interaction in 1D and confirming the usefulness of this tool, the decision was made to proceed with implementing a 3D version of the \textit{Rhodotron Simulation}.
Although complete refactoring of the software was necessary, this upgrade was crucial for implementation of \eB interaction.
The refactoring effort included proper implementation of \textit{OOP}, details of which can be seen in \fromfigsix{e_class}{bunch_class}{gun_class}{sim_class}{rhodosim_class}{rf_class} of \fromapp{intermediate_codes}.

@@

Simulating in 3D had one other benefit; it was now possible to visualize the results by rendering the interaction data. 
For this purpose, \textit{gnuplot} was integrated into \textit{Rhodotron Simulation} to produce 2D visualization of acceleration plane. 
Rendered results could be stored as \textit{gif} animations. 
Two of such renders can be seen in \fromfig{example_gnuplot_renders}.
\iffalse \begin{figure}[H]
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \centering
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/5ns_gnuplot.png}
        \caption*{Unsyncronized \e gun period, $T_g = 5ns$}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/9_3ns_gnuplot.png}
        \caption*{Syncronized \e gun period, $T_g = 9.3ns$}
    \end{subfigure}
    \caption{Example \textit{gnuplot} renders of \textit{Rhodotron Simulation} $P=40$kW, $f=107.5$MHz.}
    \label{fig:example_gnuplot_renders}
\end{figure} \fi
\begin{figure}[H]
    \centering
    \subfigure[\centering Unsyncronized \e gun period, $T_g = 5ns$]{{\includegraphics[width=0.85\textwidth]{../../../figures/rhodoSim/5ns_gnuplot.png}}}%
    \vspace{15pt}
    \centering
    \subfigure[\centering Syncronized \e gun period, $T_g = 9.3ns$]{{\includegraphics[width=0.85\textwidth]{../../../figures/rhodoSim/9_3ns_gnuplot.png}}}%
    \vspace{15pt}
    \caption{Example \textit{gnuplot} renders of \textit{Rhodotron Simulation}, $P=40$kW, $f=107.5$MHz.} 
    \label{fig:example_gnuplot_renders}
\end{figure}

\subsection{Acceleration in Magnetic Field}
An issue regarding the \eB interaction became apparent when energy gain during these interactions was observed.
A setup simulation was implemented in which a bunch of 100\e at $1$MeV was fired into a uniform magnetic field of $0.1$T located in $x>0.05$m.
Initial results at $dt=0.01$ns proved the suspicion of \eB interaction being broken. However, the energy gain would decrease tremendously as $dt$ decreased.
\iffalse \begin{figure}
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \centering
    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/mag_lf_001dt.png}
        \caption*{$dt=10^{-2}$ns, $\Delta E=2.783$MeV}
    \end{subfigure}
    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/mag_lf_00001dt.png}
        \caption*{$dt=10^{-4}$ns, $\Delta E=0.011$MeV}
    \end{subfigure}
    \caption{Energy gain of $1$MeV bunch in \textbf{B}=$0.1$T}
    \label{fig:mag_lf_render}
\end{figure} \fi
\begin{figure}
    \centering
    \subfigure[\centering $dt=10^{-2}$ns, $\Delta E=2.783$MeV]{{\includegraphics[width=0.8\textwidth]{../../../figures/rhodoSim/mag_lf_001dt.png} }}%
    \vspace{15pt}
    \subfigure[\centering $dt=10^{-4}$ns, $\Delta E=0.011$MeV]{{\includegraphics[width=0.8\textwidth]{../../../figures/rhodoSim/mag_lf_00001dt.png} }}%
    \vspace{15pt}
    \caption{Energy gain of $1$MeV bunch in \textbf{B}=$0.1$T.}
    \label{fig:mag_lf_render}
\end{figure}
Decreasing $dt$ would be the best way to increase accuracy of the results; however, this is not sustainable due to computing power limitations. 
Until this point, \textit{Rhodotron Simulation} have been using \fromsec{leapfrog} for \eEM interactions. 
To test newer approaches, two additional version of  \eEM interaction that are using \fromsec{rungekutta} were added.
\subsubsection{RK4-1}
First approach for integrating \fromsec{rungekutta} into \eEM interaction was to calculate $\vecthreeBF{a}_{E}$ and $\vecthreeBF{a}_{B}$ from \fromeq{acc_of_E_and_B} using \textit{RK4}.
After $\vecthreeBF{a}_{EM} = \vecthreeBF{a}_{E} + \vecthreeBF{a}_{B}$ was calculated, \e would move and accelerate using the \textit{Leap-frog} method. 
The idea was to produce more refined interaction results, leading to improved accuracy especially in \eB.
RF field was kept static during the RK4 computation, due to ongoing \textit{multithreading} implementation efforts. 
The implementation can be found in \fromfigs{rk1_B}{rk1_EM}.
\subsubsection{RK4-2}
Following the implementation of \textit{RK4-1}, revisions were made to the integration method for \textit{RK4} to replace \textit{Leap-frog}.
Instead of calculating $\vecthreeBF{a}_{EM}$ using \textit{RK4}, $\vecthreeBF{r}$ and $\vecthreeBF{v}$ would be determined directly.

These three methods were then tested in the same setup as \fromfig{mag_lf_render}. The results can be found in \fromfig{lf_rk1_rk2_comparison} of \fromapp{data_graph}.
\textit{RK4-1} was decided to be abandoned as it produced the same accuracy in twice the simulation time of \textit{RK4-2}.
\clearpage
\iffalse \begin{figure}
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \centering
    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/mag_lf_rk2_dt-E.png}
        \caption*{$dt$ vs $\Delta E$}
    \end{subfigure}

    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/mag_lf_rk2_dt-Tsim.png}
        \caption*{$dt$ vs $T_{sim}$}
    \end{subfigure}
    \caption{Comparing Leap-frog, RK4-2 performance on \eB interaction $E_{in}=1$MeV, \textbf{B}=$0.1$T, $t_{end}=5$ns}
    \label{fig:mag_lf_rk2_comparison}
\end{figure} \fi
\begin{figure}[h]
    \centering
    \subfigure[\centering $dt$ vs $\Delta E$]{{\includegraphics[width=0.8\textwidth]{../../../figures/analiz/mag_lf_rk2_dt-E.png} }}%
    \qquad\subfigure[\centering $dt$ vs $T_{sim}$]{{\includegraphics[width=0.8\textwidth]{../../../figures/analiz/mag_lf_rk2_dt-Tsim.png} }}%
    \vspace{20pt}
    \caption{Comparing Leap-frog, RK4-2 performance on \eB interaction $E_{in}=1$MeV, \textbf{B}=$0.1$T, $t_{end}=5$ns.}
    \label{fig:mag_lf_rk2_comparison}
\end{figure}
More rigorous testing was done with \textit{Leap-frog} and \textit{RK4-2} however. 
Still using the setup in \fromfig{mag_lf_render}, each $dt$ configuration was simulated 10 times, calculating average and standard deviations afterwards.

Example renders from these tests can be observed in \fromfig{mag_lf_rk2_comparison} and
the data from these tests can be found in \fromtabs{lf_mag_table}{rk2_mag_table} of \fromapp{data_graph}.
To investigate the data further, one can define a performance measurement, $F$ as
\vspace{-10pt}\begin{eqnarray}
    F &\propto& 1/T  , \\
    F &\propto& 1/\Delta E . 
\vspace{-10pt}\end{eqnarray}
When $dt=10^{-5}$ ns was taken as reference point due to providing a good balance of accuracy and computational intensity, 
relative performance can be calculated as
\vspace{-10pt}\begin{eqnarray} \label{eq:f_lf_rk_mag_1}
    \Delta E_{LF} &=& 110 \times 10^{-5} MeV,\\
    \Delta E_{RK}^1 &=& 64 \times 10^{-5} MeV,\\
    T_{LF}^1 &=& 9.44 \pm 0.03 s, \\
    T_{RK}^1 &=& 21.33 \pm 0.02 s, \\
    \Delta E_{LF} \times T_{LF} &=& 104 \times 10^{-4} \pm 10^{-4} MeV s, \\ %0.003
    \Delta E_{RK} \times T_{RK} &=& 137 \times 10^{-4} \pm 2\times 10^{-4}  MeV s, \\ %0.001
    F_{LF}/F_{RK} &=& \frac{\Delta E_{RK} \times T_{RK} }{\Delta E_{LF} \times T_{LF}^1} = 1.32 \pm 0.01 .
\vspace{-10pt}\end{eqnarray}
Also, another relative performance can be obsered from the data as
\vspace{-10pt}\begin{eqnarray} \label{eq:f_lf_rk_mag_2}
    \Delta E_{LF} (dt=3 \times 10^{-5}) \approx \Delta E_{RK} (dt=5 \times 10^{-5}) &\approx& 32.5 \times 10^{-4} MeV, \\
    T_{LF}(dt=3 \times 10^{-5})  &=& 3.18 \pm 0.02 s, \\ %0.0032
    T_{RK}(dt=5 \times 10^{-5})  &=& 4.30 \pm 0.01 s, \\ %0.0022
    F_{LF}/F_{RK} = \frac{T_{RK}(dt=5 \times 10^{-5})}{T_{LF}(dt=3 \times 10^{-5})} &=& 1.4 \pm 0.1.
\vspace{-10pt}\end{eqnarray}
\clearpage
Uncertainty of \fromeq{f_lf_rk_mag_2} was taken high due to the approximation. 
After combining \fromeqs{f_lf_rk_mag_1}{f_lf_rk_mag_2}, $F$ can be calculated as
\vspace{-10pt}\begin{equation}
    F_{LF}/F_{RK} = 1.36 \pm 0.05.
\vspace{-10pt}\end{equation}
Therefore, \textit{Leap-frog} was found to be the better choice as it provided with $1.36 \pm 0.05$ times accuracy in \eB interactions at a given time with respect to \textit{RK4}. However,
\textit{RK4} was promising in situations where decreasing the stepsize, $dt$, is not viable. Therefore both were integrated into \textit{Rhodotron Simulation} for the user to decide.
\textit{RK4-2} renders from these test can be found in \fromfig{mag_rk_render} of \fromapp{example_simulation_runs}.

\subsection{Acceleration in Electric Field}
After the accuracy concerns regarding the \eB interaction were raised, it was decided to test \eE and compare the performance of \textit{Leap-frog} and \textit{RK4}.

@@

\textit{Leap-frog} provides $2.4$ times and $2.0$ times less overacceleration per simulation time than 
\textit{RK4} in parallel and perpendicular electric fields respectably.
This results can be tested further with obsering from the data (see \fromtabf{lf_statE_table}{rk2_statE_table}{lf_statE90_table}{rk2_statE90_table} of \fromapp{data_graph})
and calculating other relative performances as
\vspace{-10pt}\begin{eqnarray} \label{eq:f_lf_rk_3}
    \Delta E_{LF}^1 (dt=2 \times 10^{-5}) = \Delta E_{RK}^1 (dt=3 \times 10^{-5}) &=& 45 \times 10^{-6} MeV, \\
    T_{LF}^1(dt=2 \times 10^{-5})  &=& 6.23 \pm 0.02 s, \\ %0.0032
    T_{RK}^1(dt=3 \times 10^{-5})  &=& 13.40 \pm 0.03 s, \\ %0.0022
    F_{LF}^1/F_{RK}^1 = \frac{T_{RK}^1(dt=3 \times 10^{-5})}{T_{LF}^1(dt=2 \times 10^{-5})} &=& 2.15 \pm 0.02,
\vspace{-10pt}\end{eqnarray}
and
\vspace{-10pt}\begin{eqnarray} \label{eq:f_lf_rk_4}
    \Delta E_{LF}^2 (dt=3 \times 10^{-5}) \approx \Delta E_{RK}^2 (dt=5 \times 10^{-5}) &\approx& 470 \times 10^{-6} MeV, \\
    T_{LF}^2(dt=3 \times 10^{-5})  &=& 3.62, \\ %0.0055
    T_{RK}^2(dt=5 \times 10^{-5})  &=& 7.00, \\ %0.0071 -> 0,0126
    F_{LF}^2/F_{RK}^2 = \frac{T_{RK}^2(dt=5 \times 10^{-5})}{T_{LF}^2(dt=3 \times 10^{-5})} &=& 1.9 \pm 0.1.
\vspace{-10pt}\end{eqnarray}
The last uncertainty was taken high due to the approximation of $467 \approx 471$. After combining \fromeqsf{f_lf_rk_1}{f_lf_rk_2}{f_lf_rk_3}{f_lf_rk_4},
the performance of \textit{Leap-frog} relative to \textit{RK4} in \eE interaction can be calculated as
\vspace{-10pt}\begin{equation}\label{eq:lf_rk_eE_relative_performance}
    F_{LF}/F_{RK} = 2.11 \pm 0.03.
\vspace{-10pt}\end{equation}
Therefore, it can be concluded that \textit{Leap-frog} outperforms \textit{RK4} with the relative performance of $2.11 \pm 0.03$ in \eE interactions with static uniform $\vecthreeBF{E}$ field.

\subsection{Multithreading}
As already mentioned before, multithreading implementation efforts were ongoing since right after the start of this project. 
There have been a number of different approaches for implementation. 
First implemenation was done when the \textit{Rhodotron Simulation} was only capable of 1D simulations.
After the sizable refactoring done for 3D capabilities, this implementation was obselete.

For the current version, a main thread that spawns and manages several other worker threads would be used as can be seen in \fromfig{multh_illustration}.
The UI-Console thread would handle incoming and outgoing communication, notifying the user about the status of simulation 
(see \fromfig{console_output_running} in \fromapp{example_simulation_runs} for example console notification, \fromfig{ui_thread} in \fromapp{intermediate_codes} for implemenation)
or communucating with the \textit{GUI} that will be disscussed in the next section.

@@

        # SIM CONFIGURATION 
        starttime=0
        endtime=10
        dt=0.0000010000
        epath=xy/rf.dat
        bpath=xy/magnet.dat
        cpath=xy/settings.dat
        ppath=xy/paths/
        multh=1
        thcount=10     
    \end{minted}
\caption{An example \textit{config.in} file.}
\label{fig:config_file}
\end{figure}

>>>>

\subsection{$L_{out}$ Optimization For Single \e} \label{sec:lout_sweep}
Initial step of improving POC towards Rhodotron Simulation was to implement $L_{out}$ optimizations to help optimazing magnet designs, as discussed in \fromsec{parameter_sweep}.
First approach was to hang the \e outside of the cavity for $t_{out} = L_{out}/v$, then inject it back to the cavity with reversed $\vecthreeBF{v}$. One can sweep the $t_{out}$ parameter to find the optimal value.
This simple implementation can be found in \fromfig{lout_opt_single_e} of \fromapp{intermediate_codes}.

@@

\subsection{$\phi_{lag}$ Optimization for Bunches} \label{sec:philag_sweep}
After successfully accelerating single \e, particle bunches were implemented to approximate a real \e gun. 
They were modeled as $N$ electrons fired from an \e gun at even time intervals. This approach was taken because the amount of time gun fires, defined as Gun Active Time, $t_g$, is a crucial part of pulsed \e gun design.
\clearpage
Addition of bunches would immediately proven useful when finding optimal phase lag for bunch entry.
$\phi_{lag}$ for a bunch was defined as the RF phase when the first \e of the bunch has entered the cavity and it defines the starting time of the current pass.
To use the parameter sweep method, as used in $L_{out}$ optimization, relevant bunch characteristics are defined as
\vspace{-10pt}\begin{itemize}
    \item $\mu E$: Average energy,
    \item $E_{rms}$: Root mean square of energy,
    \item $R_{rms}$: Root mean square of \e positions.
\end{itemize}

Optimal $\phi_{lag}$ would produce maximum $\mu E$ while minimizing $E_{rms}$ \& $R_{rms}$. For the first pass, 
$E_{rms}$ and $R_{rms}$ would be vaguely dependent of each other; therefore, initial implementation of $\phi_{lag}$ sweep was based on minimizing $E_{rms}$ during simulation (see \fromfig{phlag_opt}). 
For $\mu E$ considerations, data from the software would be analyzed either manually or by using external tools such as ROOT framework. 
\vspace{20pt}
\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
        int phase_opt(int phase_sweep_range){
            double minrms = 1;
            int opt_phase;
            for(int RFphase = -phase_sweep_range; RFphase <= phase_sweep_range; RFphase++){
              Bunch bunch1(RFphase);
              double t1 = 0;
              bunch1.bunch_gecis_t(t1);
              bunch1.reset_pos();

              if( bunch1.E_rms() < minrms ){
                minrms = bunch1.E_rms();
                opt_phase = RFphase;
              }
            }
            return opt_phase;
        }
    \end{minted}
    \vspace{20pt}
    \caption{$\phi_{lag}$ Optimization For Initial Bunch Design.}
    \label{fig:phlag_opt}
\end{figure}
\clearpage
Since $\phi_{lag}$ is relatively easy to change after production, another version of \fromfig{phlag_opt} that was modified for given magnet design parameters was also implemented ({see \fromfig{phlag_opt_n_pass} in \fromapp{intermediate_codes}). 
This version can be useful for optimizing $\phi_{lag}$ in case of production issues in magnets.

@@

\subsection{Simulation in 3D}
After successfully implementing \eE interaction in 1D and confirming the usefulness of this tool, the decision was made to proceed with implementing a 3D version of the Rhodotron Simulation.
Although complete refactoring of the software was necessary, this upgrade was crucial for implementation of \eB interaction.
The refactoring effort included proper implementation of OOP, details of which can be seen in \fromfigsix{e_class}{bunch_class}{gun_class}{sim_class}{rhodosim_class}{rf_class} of \fromapp{intermediate_codes}.

@@

Simulating in 3D had one other benefit; it was now possible to visualize the results by rendering the interaction data. 
For this purpose, gnuplot was integrated into Rhodotron Simulation to produce 2D visualization of acceleration plane. 
Rendered results could be stored as gif animations. 
Two of such renders can be seen in \fromfig{example_gnuplot_renders}.
\iffalse \begin{figure}[H]
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \centering
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/5ns_gnuplot.png}
        \caption*{Unsyncronized \e gun period, $T_g = 5ns$}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/9_3ns_gnuplot.png}
        \caption*{Syncronized \e gun period, $T_g = 9.3ns$}
    \end{subfigure}
    \caption{Example gnuplot renders of Rhodotron Simulation $P=40$kW, $f=107.5$MHz.}
    \label{fig:example_gnuplot_renders}
\end{figure} \fi
\begin{figure}[H]
    \centering
    \subfigure[\centering Unsyncronized \e gun period, $T_g = 5ns$]{{\includegraphics[width=0.85\textwidth]{../../../figures/rhodoSim/5ns_gnuplot.png}}}%
    \vspace{15pt}
    \centering
    \subfigure[\centering Syncronized \e gun period, $T_g = 9.3ns$]{{\includegraphics[width=0.85\textwidth]{../../../figures/rhodoSim/9_3ns_gnuplot.png}}}%
    \vspace{15pt}
    \caption{Example gnuplot renders of Rhodotron Simulation, $P=40$kW, $f=107.5$MHz.} 
    \label{fig:example_gnuplot_renders}
\end{figure}

\subsection{Acceleration in Magnetic Field}
An issue regarding the \eB interaction became apparent when energy gain during these interactions was observed.
A setup simulation was implemented in which a bunch of 100\e at $1$MeV was fired into a uniform magnetic field of $0.1$T located in $x>0.05$m.
Initial results at $dt=0.01$ns proved the suspicion of \eB interaction being broken. However, the energy gain would decrease tremendously as $dt$ decreased.
\iffalse \begin{figure}
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \centering
    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/mag_lf_001dt.png}
        \caption*{$dt=10^{-2}$ns, $\Delta E=2.783$MeV}
    \end{subfigure}
    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/mag_lf_00001dt.png}
        \caption*{$dt=10^{-4}$ns, $\Delta E=0.011$MeV}
    \end{subfigure}
    \caption{Energy gain of $1$MeV bunch in \textbf{B}=$0.1$T}
    \label{fig:mag_lf_render}
\end{figure} \fi
\begin{figure}
    \centering
    \subfigure[\centering $dt=10^{-2}$ns, $\Delta E=2.783$MeV]{{\includegraphics[width=0.8\textwidth]{../../../figures/rhodoSim/mag_lf_001dt.png} }}%
    \vspace{15pt}
    \subfigure[\centering $dt=10^{-4}$ns, $\Delta E=0.011$MeV]{{\includegraphics[width=0.8\textwidth]{../../../figures/rhodoSim/mag_lf_00001dt.png} }}%
    \vspace{15pt}
    \caption{Energy gain of $1$MeV bunch in \textbf{B}=$0.1$T.}
    \label{fig:mag_lf_render}
\end{figure}
Decreasing $dt$ would be the best way to increase accuracy of the results; however, this is not sustainable due to computing power limitations. 
Until this point, Rhodotron Simulation have been using \fromsec{leapfrog} for \eEM interactions. 
To test newer approaches, two additional version of  \eEM interaction that are using \fromsec{rungekutta} were added.
\subsubsection{RK4-1}
First approach for integrating \fromsec{rungekutta} into \eEM interaction was to calculate $\vecthreeBF{a}_{E}$ and $\vecthreeBF{a}_{B}$ from \fromeq{acc_of_E_and_B} using RK4.
After $\vecthreeBF{a}_{EM} = \vecthreeBF{a}_{E} + \vecthreeBF{a}_{B}$ was calculated, \e would move and accelerate using the Leap-frog method. 
The idea was to produce more refined interaction results, leading to improved accuracy especially in \eB.
RF field was kept static during the RK4 computation, due to ongoing multithreading implementation efforts. 
The implementation can be found in \fromfigs{rk1_B}{rk1_EM}.
\subsubsection{RK4-2}
Following the implementation of RK4-1, revisions were made to the integration method for RK4 to replace Leap-frog.
Instead of calculating $\vecthreeBF{a}_{EM}$ using RK4, $\vecthreeBF{r}$ and $\vecthreeBF{v}$ would be determined directly.

These three methods were then tested in the same setup as \fromfig{mag_lf_render}. The results can be found in \fromfig{lf_rk1_rk2_comparison} of \fromapp{data_graph}.
RK4-1 was decided to be abandoned as it produced the same accuracy in twice the simulation time of RK4-2.
\clearpage
\iffalse \begin{figure}
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \centering
    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/mag_lf_rk2_dt-E.png}
        \caption*{$dt$ vs $\Delta E$}
    \end{subfigure}

    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/mag_lf_rk2_dt-Tsim.png}
        \caption*{$dt$ vs $T_{sim}$}
    \end{subfigure}
    \caption{Comparing Leap-frog, RK4-2 performance on \eB interaction $E_{in}=1$MeV, \textbf{B}=$0.1$T, $t_{end}=5$ns}
    \label{fig:mag_lf_rk2_comparison}
\end{figure} \fi
\begin{figure}[h]
    \centering
    \subfigure[\centering $dt$ vs $\Delta E$]{{\includegraphics[width=0.8\textwidth]{../../../figures/analiz/mag_lf_rk2_dt-E.png} }}%
    \qquad\subfigure[\centering $dt$ vs $T_{sim}$]{{\includegraphics[width=0.8\textwidth]{../../../figures/analiz/mag_lf_rk2_dt-Tsim.png} }}%
    \vspace{20pt}
    \caption{Comparing Leap-frog, RK4-2 performance on \eB interaction $E_{in}=1$MeV, \textbf{B}=$0.1$T, $t_{end}=5$ns.}
    \label{fig:mag_lf_rk2_comparison}
\end{figure}
More rigorous testing was done with Leap-frog and RK4-2 however. 
Still using the setup in \fromfig{mag_lf_render}, each $dt$ configuration was simulated 10 times, calculating average and standard deviations afterwards.

Example renders from these tests can be observed in \fromfig{mag_lf_rk2_comparison} and
the data from these tests can be found in \fromtabs{lf_mag_table}{rk2_mag_table} of \fromapp{data_graph}.
To investigate the data further, one can define a performance measurement, $F$ as
\vspace{-10pt}\begin{eqnarray}
    F &\propto& 1/T  , \\
    F &\propto& 1/\Delta E . 
\vspace{-10pt}\end{eqnarray}
When $dt=10^{-5}$ ns was taken as reference point due to providing a good balance of accuracy and computational intensity, 
relative performance can be calculated as
\vspace{-10pt}\begin{eqnarray} \label{eq:f_lf_rk_mag_1}
    \Delta E_{LF} &=& 110 \times 10^{-5} MeV,\\
    \Delta E_{RK}^1 &=& 64 \times 10^{-5} MeV,\\
    T_{LF}^1 &=& 9.44 \pm 0.03 s, \\
    T_{RK}^1 &=& 21.33 \pm 0.02 s, \\
    \Delta E_{LF} \times T_{LF} &=& 104 \times 10^{-4} \pm 10^{-4} MeV s, \\ %0.003
    \Delta E_{RK} \times T_{RK} &=& 137 \times 10^{-4} \pm 2\times 10^{-4}  MeV s, \\ %0.001
    F_{LF}/F_{RK} &=& \frac{\Delta E_{RK} \times T_{RK} }{\Delta E_{LF} \times T_{LF}^1} = 1.32 \pm 0.01 .
\vspace{-10pt}\end{eqnarray}
Also, another relative performance can be obsered from the data as
\vspace{-10pt}\begin{eqnarray} \label{eq:f_lf_rk_mag_2}
    \Delta E_{LF} (dt=3 \times 10^{-5}) \approx \Delta E_{RK} (dt=5 \times 10^{-5}) &\approx& 32.5 \times 10^{-4} MeV, \\
    T_{LF}(dt=3 \times 10^{-5})  &=& 3.18 \pm 0.02 s, \\ %0.0032
    T_{RK}(dt=5 \times 10^{-5})  &=& 4.30 \pm 0.01 s, \\ %0.0022
    F_{LF}/F_{RK} = \frac{T_{RK}(dt=5 \times 10^{-5})}{T_{LF}(dt=3 \times 10^{-5})} &=& 1.4 \pm 0.1.
\vspace{-10pt}\end{eqnarray}
\clearpage
Uncertainty of \fromeq{f_lf_rk_mag_2} was taken high due to the approximation. 
After combining \fromeqs{f_lf_rk_mag_1}{f_lf_rk_mag_2}, $F$ can be calculated as
\vspace{-10pt}\begin{equation}
    F_{LF}/F_{RK} = 1.36 \pm 0.05.
\vspace{-10pt}\end{equation}
Therefore, Leap-frog was found to be the better choice as it provided with $1.36 \pm 0.05$ times accuracy in \eB interactions at a given time with respect to RK4. However,
RK4 was promising in situations where decreasing the stepsize, $dt$, is not viable. Therefore both were integrated into Rhodotron Simulation for the user to decide.
RK4-2 renders from these test can be found in \fromfig{mag_rk_render} of \fromapp{example_simulation_runs}.

\subsection{Acceleration in Electric Field}
After the accuracy concerns regarding the \eB interaction were raised, it was decided to test \eE and compare the performance of Leap-frog and RK4.

@@

Leap-frog provides $2.4$ times and $2.0$ times less overacceleration per simulation time than 
RK4 in parallel and perpendicular electric fields respectably.
This results can be tested further with obsering from the data (see \fromtabf{lf_statE_table}{rk2_statE_table}{lf_statE90_table}{rk2_statE90_table} of \fromapp{data_graph})
and calculating other relative performances as
\vspace{-10pt}\begin{eqnarray} \label{eq:f_lf_rk_3}
    \Delta E_{LF}^1 (dt=2 \times 10^{-5}) = \Delta E_{RK}^1 (dt=3 \times 10^{-5}) &=& 45 \times 10^{-6} MeV, \\
    T_{LF}^1(dt=2 \times 10^{-5})  &=& 6.23 \pm 0.02 s, \\ %0.0032
    T_{RK}^1(dt=3 \times 10^{-5})  &=& 13.40 \pm 0.03 s, \\ %0.0022
    F_{LF}^1/F_{RK}^1 = \frac{T_{RK}^1(dt=3 \times 10^{-5})}{T_{LF}^1(dt=2 \times 10^{-5})} &=& 2.15 \pm 0.02,
\vspace{-10pt}\end{eqnarray}
and
\vspace{-10pt}\begin{eqnarray} \label{eq:f_lf_rk_4}
    \Delta E_{LF}^2 (dt=3 \times 10^{-5}) \approx \Delta E_{RK}^2 (dt=5 \times 10^{-5}) &\approx& 470 \times 10^{-6} MeV, \\
    T_{LF}^2(dt=3 \times 10^{-5})  &=& 3.62, \\ %0.0055
    T_{RK}^2(dt=5 \times 10^{-5})  &=& 7.00, \\ %0.0071 -> 0,0126
    F_{LF}^2/F_{RK}^2 = \frac{T_{RK}^2(dt=5 \times 10^{-5})}{T_{LF}^2(dt=3 \times 10^{-5})} &=& 1.9 \pm 0.1.
\vspace{-10pt}\end{eqnarray}
The last uncertainty was taken high due to the approximation of $467 \approx 471$. After combining \fromeqsf{f_lf_rk_1}{f_lf_rk_2}{f_lf_rk_3}{f_lf_rk_4},
the performance of Leap-frog relative to RK4 in \eE interaction can be calculated as
\vspace{-10pt}\begin{equation}\label{eq:lf_rk_eE_relative_performance}
    F_{LF}/F_{RK} = 2.11 \pm 0.03.
\vspace{-10pt}\end{equation}
Therefore, it can be concluded that Leap-frog outperforms RK4 with the relative performance of $2.11 \pm 0.03$ in \eE interactions with static uniform $\vecthreeBF{E}$ field.

\subsection{Multithreading}
As already mentioned before, multithreading implementation efforts were ongoing since right after the start of this project. 
There have been a number of different approaches for implementation. 
First implemenation was done when the Rhodotron Simulation was only capable of 1D simulations.
After the sizable refactoring done for 3D capabilities, this implementation was obselete.

For the current version, a main thread that spawns and manages several other worker threads would be used as can be seen in \fromfig{multh_illustration}.
The UI-Console thread would handle incoming and outgoing communication, notifying the user about the status of simulation 
(see \fromfig{console_output_running} in \fromapp{example_simulation_runs} for example console notification, \fromfig{ui_thread} in \fromapp{intermediate_codes} for implemenation)
or communucating with the GUI that will be disscussed in the next section.

@@

        # SIM CONFIGURATION 
        starttime=0
        endtime=10
        dt=0.0000010000
        epath=xy/rf.dat
        bpath=xy/magnet.dat
        cpath=xy/settings.dat
        ppath=xy/paths/
        multh=1
        thcount=10     
    \end{minted}
\caption{An example config.in file.}
\label{fig:config_file}
\end{figure}
